True Stellar Obliquity Distribution of Exoplanets

### 2022-11-19 src/tex/ms.tex
[showyourwork] overleaf sync

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

@@

% Abstract with filler text
\begin{abstract}
We build a hierarchical Bayesian framework to constrain true stellar obliquity ($\psi$) distribution in exoplanetary systems.
Using the sky-projected stellar obliquities ($\lambda$) of a sample of 151 exoplanets, we infer their $\cos{\psi}$ distribution using a four-parameter Beta mixture model with the assumption of a uniform cosine of stellar inclination ($i_\star$) distribution for each planet. We find $\cos{\psi}$ has a significant pile-up near 1, suggesting the majority of exoplanetary systems are well aligned. For misaligned systems, $\cos{\psi}$ is nearly uniformly distributed between $-0.75$ to $0.75$ with no significant pipe-up near $\cos{\psi} = 0$, suggesting the stellar spin axis is randomly distributed around the planet's orbital axis.
Our Bayesian framework could reproduce the excess of perpendicular planets observed in systems with constrained $i_\star$. However, we find the sample of misaligned systems could be biased because of the small sample size and that misaligned systems are often found around hot stars, whereas the gravity-darkening method, a technique commonly used to constrain the inclinations of hot stars ($T_{\rm eff} > 7000$K), is only sensitive to detect perpendicular planets. 
\end{abstract}

We build a hierarchical Bayesian framework to constrain true stellar obliquity ($\psi$) distribution in exoplanetary systems.

@@

Stellar obliquity $\psi$ can be calculated from $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ as the following:
\begin{equation}
    \cos{\psi} = \sin{i_\star}\sin{i_{\rm orb}}\cos{\lambda} + \cos{i_\star}\cos{i_{\rm orb}},
\end{equation}
where $i_{\rm orb}$ and $i_\star$ are the projections of $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ along the observer's line of sight, and $\lambda$ is the sky-projected stellar obliquity. A full derivation can be found in, e.g., \cite{Fabrycky09} Equation (1)--(10).
When setting $i_{\rm orb} = 90\degr$, $\cos{\psi} = \sin{i_\star}\cos{\lambda}$.

@@

\section{Stellar Obliquity Distribution} \label{sec:hbm}
We model the whole population stellar obliquity distribution and also the sub-population with $i_\star$ constrained. See Figure\ref{fig:hbm} as an illustration.

\begin{figure}[ht!]
    \script{hbm.py}
    \includegraphics[width=\linewidth]{figures/hbm.pdf}
    \caption{Hierarchical Bayesian framework to infer stellar obliquity distribution described by hyperparameters $\theta$ from individual system's stellar obliquity $\psi_n$. $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$.}
    \label{fig:hbm}
\end{figure}

We use a four-parameter Beta distribution. The framework is shown in Figure~\ref{fig:hbm}.

@@

\begin{figure*}[ht!]
    \script{psi_dist.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/psi_dist.pdf}
        \caption{$\cos{\psi}$ distributions.}
        \label{fig:psi_dist}
    \end{centering}
\end{figure*}

@@

@@

We removed 7/21 planets with host star effective temperatures greater than 7000 K and redo the analysis. Interestingly, we find the results do not change. See Figure~\ref{fig:teteff_cutff}.

>>>>

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage{xspace}

\newcommand{\numistar}{65\xspace}
\newcommand{\numall}{161\xspace}

@@

% Abstract with filler text
\begin{abstract}
We build a hierarchical Bayesian framework to constrain true stellar obliquity ($\psi$) distribution in exoplanetary systems.
Using the sky-projected stellar obliquities ($\lambda$) of all measured systems, we infer their $\cos{\psi}$ distribution using a four-parameter Beta mixture model with the assumption of a uniform cosine of stellar inclination ($i_\star$) distribution for each planet. We find $\cos{\psi}$ has a significant pile-up near 1, suggesting the majority of exoplanetary systems are well aligned. For misaligned systems, $\cos{\psi}$ is nearly uniformly distributed between $-0.75$ to $0.75$ with no significant pipe-up near $\cos{\psi} = 0$, suggesting the stellar spin axis is randomly distributed around the planet's orbital axis.
Our Bayesian framework could reproduce the excess of perpendicular planets observed in systems with constrained $i_\star$. However, we find the sample of misaligned systems could be biased because of the small sample size and that misaligned systems are often found around hot stars, whereas the gravity-darkening method, a technique commonly used to constrain the inclinations of hot stars ($T_{\rm eff} > 7000$K), is only sensitive to detect perpendicular planets. 
\end{abstract}

@@

Stellar obliquity $\psi$ can be calculated from $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ as the following:
\begin{equation}
    \cos{\psi} = \sin{i_\star}\sin{i_{\rm orb}}\cos{\lambda} + \cos{i_\star}\cos{i_{\rm orb}},
\end{equation}
where $i_{\rm orb}$ and $i_\star$ are the projections of $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ along the observer's line of sight, and $\lambda$ is the sky-projected stellar obliquity. A full derivation can be found in, e.g., \cite{Fabrycky09} Equation (1)--(10).
For transiting planet systems, $i_{\rm orb} \approx 90\degr$, and thus $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.

@@

\section{Stellar Obliquity Distribution} \label{sec:hbm}
We model the whole population stellar obliquity distribution and also the sub-population with $i_\star$ constrained. See Figure\ref{fig:graph} as an illustration.

\begin{figure}[ht!]
    \script{graph.py}
    \includegraphics[width=\linewidth]{figures/graph.pdf}
    \caption{Hierarchical Bayesian framework to infer stellar obliquity distribution described by hyperparameters $\theta$ from individual system's stellar obliquity $\psi_n$. $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$.}
    \label{fig:graph}
\end{figure}

We use a four-parameter Beta distribution.

@@

Although the distribution of $\psi$ looks very similar for the $i_\star$ sample using constrained or random priors, the uncertainty on each $\psi$ is much tighter.

\begin{figure*}[ht!]
    \script{psi_dist.py}
    \begin{centering}
        \includegraphics{figures/psi_dist.pdf}
        \caption{$\cos{\psi}$ distributions.}
        \label{fig:psi_dist}
    \end{centering}
\end{figure*}

@@

\begin{figure*}[ht!]
    \script{lam_dist.py}
    \begin{centering}
        \includegraphics{figures/lam_dist.pdf}
        \caption{Sky-projected stellar obliquity ($\lambda$) distributions of \numistar systems with constrained $i_\star$ and all \numall exoplanetary systems that have measured sky-projected stellar obliquities.}
        \label{fig:lam_dist}
    \end{centering}
\end{figure*}

@@

We removed 7/21 planets with host star effective temperatures greater than 7000 K and redo the analysis. Interestingly, we find the results do not change. See Figure~\ref{fig:teff_cutff}.
